#Code for reading in the dataset and/or processing the data
```{r message=FALSE}
rawdata <- read.csv("activity.csv", header = TRUE, na.strings = "NA")
#convert date column from factor into date
rawdata$date <- as.Date(rawdata$date, "%Y-%m-%d")
```
#Histogram of the total number of steps taken each day
```{r message=FALSE}
library(ggplot2)
qplot(steps, data=rawdata, xlab = "Number of Steps", ylab = "Frequency",main="Frequency of Steps")
```
#Mean and median number of steps taken each day
```{r message=FALSE}
library(dplyr)
#summarise
summariseddata <- rawdata %>%
        group_by(date) %>%
        summarise(median = median(steps),average = mean(steps))

print(summariseddata)

```
#Time series plot of the average number of steps taken
```{r message=FALSE}
#plot the time series
gplot <- ggplot(data = summariseddata, aes(date, average))
gplot+geom_line() + geom_smooth() + labs(x= "Date", y="Average Number of Steps")
```

#The 5-minute interval that, on average, contains the maximum number of steps
```{r message=FALSE}
#group by interval
sumdata <- rawdata[,c(3,1)] %>%
        group_by(interval)
#remove NAs
isna <- is.na(sumdata$steps)
sumdata <- sumdata[!isna,]

#find average
sumdata <- sumdata <- summarise(sumdata, avg = mean(steps))

#find maximum average and return the interval
sumdata[which.max(sumdata$avg),]
```
#Code to describe and show a strategy for imputing missing data

#Histogram of the total number of steps taken each day after missing values are imputed
#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
#All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
