---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The data is to be read in, and date is to be converted from text to date.
```{r message=FALSE}
#unzip the file
unzip("activity.zip", exdir = getwd())
rawdata <- read.csv("activity.csv", header = TRUE, na.strings = "NA")
#convert date column from factor into date
rawdata$date <- as.Date(rawdata$date, "%Y-%m-%d")


```

## What is mean total number of steps taken per day?

Calculate and report the mean and median of the total number of steps taken per day. 

```{r message=FALSE}
library(dplyr)
#summarise
summariseddata <- rawdata %>%
        group_by(date) %>%
        summarise(median = median(steps),average = mean(steps), total = sum(steps))
library(kableExtra)

summariseddata %>%
        kable() %>%
          kable_styling()
```
Make a histogram of the total number of steps taken each day. 
```{r message=FALSE}
#plot the time series
library(ggplot2)
gplot <- ggplot(data = summariseddata, aes(total))
gplot+geom_histogram(fill="skyblue", bins = 12) + labs(x= "Total Steps", y="Count")

```

## What is the average daily activity pattern?

```{r message=FALSE}
#group by interval
library(dplyr)
sumdata <- rawdata[,c(3,1)] %>%
        group_by(interval)
#remove NAs
isna <- is.na(sumdata$steps)
sumdata <- sumdata[!isna,]

#find average
sumdata <-  summarise(sumdata, avg = mean(steps))
```

```{r message=FALSE, echo=FALSE}
#find maximum average and return the interval
max <- which.max(sumdata$avg)
intv <- sumdata[which.max(sumdata$avg),1]
```

The interval with the highest average is `r max`. It comes from interval `r intv`. 

Plot the average over time.

```{r message=FALSE, echo=FALSE}
plot(sumdata$interval, sumdata$avg, type="l", xlab= "Time Interval", ylab = "Average Steps", main = "Average Number of Steps per Time Interval")
```


## Imputing missing values
The impute method is going to be the average for the interval.
```{r}

#count number of missing values
#set up a logical data frame
navals <- is.na(rawdata)
nonavals <- sum(navals[,1] == TRUE)
#count missing vals in first col (no NAs in other columns)

```

The total number of NAs are `r nonavals`. 

Plot the total steps per day.

```{r}
#Impute missing values with the mean of each interval
impdata <- rawdata %>% 
        group_by(interval) %>% 
        mutate(steps = ifelse(is.na(steps), mean(steps, na.rm=TRUE),steps))

#create table for total steps taken each day
totalsteps <- impdata %>% 
        ungroup() %>%
        group_by(date) %>%
        summarise(total_steps = mean(steps))

#plot out the histogram
impplot <- ggplot(data=totalsteps, aes(total_steps))
impplot + geom_histogram(fill="hotpink2", bins = 12)+labs(x="Total Steps", y="Count", title="Count of Daily Total Steps")

```


## Are there differences in activity patterns between weekdays and weekends?
I'm going to use the lattice package to plot this.

There are two notable trends in the graphs below: 

* the person wakes up later on the weekend and they are more active overall during the day. 

* the person has the highest activity in the morning during the weekday, as we can assume they are getting ready for work. We can consider that they have an office job as their activity peaks during lunch, work finish and dinner time.

```{r}
library(lattice)
compdata <- impdata %>%
        ungroup() %>% #ungroup the imputed dataset
        #create a column that has weekday and weekend
        mutate(day = case_when(weekdays(date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday", weekdays(date) %in% c("Saturday", "Sunday")~"weekend"))

#convert to factor
compdata$day <- as.factor(compdata$day) 
#calculate average, median and total steps
compdata <- compdata %>% 
        group_by(day, interval) %>%
        summarise(median = median(steps),average = mean(steps), total = sum(steps))



#plot the facets - reviewed total and median - not meaningful pictures
xyplot(average ~ interval | day, data = compdata, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Average Steps", main = "Average Steps by Interval")

```
